```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(gtable)
library(scales)
library(lubridate)
```

```{r}
setwd("~/GitHub/Internship")
covid_data <- read.csv('owid-covid-data.csv')
# Steven: I recommend reading directly from the internet
covid_data <- read_csv('https://covid.ourworldindata.org/data/owid-covid-data.csv')

head(covid_data)
names(covid_data)
```
```{r}
ggplot(covid_data, aes(covid_data$date, covid_data$total_cases))
# Steven: I recommend using ggplot. Also, try to choose quantities on x and y axes that will give
# a more meaningful plot
# ggplot(covid_data, aes(x=date, y=total_cases)) + geom_line(color = 'red')
```
```{r}
require(tidyverse)
# Steven: this is redundant - you already loaded tidyverse
```
```{r}
covid_data
```
filter for the columns necessary
```{r}

```
```{r}
UK_covid_data <- covid_data %>% select(location, date, total_cases,total_cases_per_million, new_cases, total_deaths, total_deaths_per_million, new_deaths, new_deaths_per_million, weekly_hosp_admissions, weekly_icu_admissions_per_million, total_tests, total_tests_per_thousand, positive_rate, total_vaccinations, total_vaccinations_per_hundred, new_vaccinations, new_vaccinations_smoothed_per_million, median_age, aged_65_older) %>% filter(location == "United Kingdom")
# Steven: UK_covid_data looks like it's the same thing as covid_data. Only need one of these
#Safia: I have removed covid_data
UK_covid_data
```

The table containing data I find to be relative and useful to the research questions I'm trying to answer are stored under the variable UK_covid_data
Steven: UK_covid_data is a dataframe. Better to call it that instead of a variable.

change the date from 'character' to 'date'
```{r}
UK_covid_data
class(UK_covid_data$date)
UK_covid_data$date<- as.Date(UK_covid_data$date)
class(UK_covid_data$date)
class(UK_covid_data$total_cases)
```
plot a graph to show the progression of total covid cases
```{r}
theme_set(theme_minimal())
ggplot(UK_covid_data, aes(date, total_cases))+
  geom_line()+
  scale_x_date(date_breaks = "2 months", date_labels = "%b%y" )+
  scale_y_continuous(labels = label_number())
# Steven: I recommend setting date_breaks = "2 month". Otherwise the lables on  axis get squashed and its hard to read.
#Safia: edited from "1 month" to "2 month"
ggsave("UK Total cases time series.png")
```
using the same code to plot new cases
```{r}
theme_set(theme_minimal())
ggplot(UK_covid_data, aes(date, new_cases))+
  geom_line()+
  scale_x_date(date_breaks = "2 month", date_labels = "%b%y" )+
  scale_y_continuous(labels = label_comma(), )+
  labs(title = "Daily new cases of Covid-19 in the UK", x= "Date", y="no. New cases")
# Steven: same thing, set date_breaks = "2 month"
#Safia: edited
ggsave("UK New cases time series .png")
```
plotting new cases with new deaths
```{r}
ggplot()+
  geom_line(aes(UK_covid_data$date, UK_covid_data$new_cases), colour= "red", show.legend = TRUE)+
  geom_line(aes(UK_covid_data$date, UK_covid_data$new_deaths), colour= "blue", show.legend = TRUE)+
  labs(title= "Reports of daily new deaths and new cases", y= "no. reports (log10)", x= "Date")+
  scale_y_log10(labels = label_comma())+
  geom_vline(xintercept = as.numeric(as.Date("2020-12-08")))+
  annotate("text", x= as.Date("2020-12-08"), y= 60000, label= "Vaccine 21+ (8Dec)", angle= 90, size= 2)+
  geom_vline(xintercept = as.numeric(as.Date("2021-06-16")))+
  annotate("text", x= as.Date("2021-06-16"), y= 60000, label= "Vaccine U21 (8Dec)", angle= 90, size= 2)
ggsave("UK new deaths and new cases time series (vaccine) time series.png")
```
when I tried changing the y axis scale to log10 to make it easier to compare the two, it introduced infinite values into the continuous y axis
Steven: Probably because log(0) = -infinity.
There appear to be two days where the number of new cases is negative. Can you explain what is happening there?
Safia:

new vaccinations time series
```{r}
vaccine<- c(if_else(UK_covid_data$date < "2021-06-16", 1, 2))
length(vaccine)<- 227
ggplot(UK_covid_data %>% filter(date>= "2020-12-08"), aes(date, new_vaccinations, colour= as.factor(vaccine)))+
  geom_point(size= 0.2)+
  geom_smooth()+
  scale_x_date(breaks = "1 week", date_labels = "%d-%m-%y")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size= 4))+
  geom_vline(xintercept = as.numeric(as.Date("2021-06-16")), linetype= 2)+
  annotate("text", x= as.Date("2021-06-16"), y= 60000, label= "Vaccine U21 (16Jun)", angle= 90, size= 2)+
  scale_y_continuous(label= label_comma())+
  labs(title = "New vaccinations in UK", y= "no. New Vaccinations", x= "Date")
ggsave("New vaccinations in the UK (scatter).png")
```

```{r}
time.points <- seq.Date(as.Date("2020-12-08"), by=1, length.out = 365)
```

 
